---
title: "Final Project"
output: 
    flexdashboard::flex_dashboard:
        orientation: columns
        vertical_layout: scroll
    html_document: default
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readxl)
library(unpivotr)
library(tidyverse)
library(tidyr)
library(htmlwidgets)
library(htmltools)
library(networkD3)
library(ggplot2)
library(ggbeeswarm)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(rworldmap)
library(hrbrthemes)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
library(tm)


```



```{r}
#read data
df = read.csv('deep_sea_corals.csv')

```

```{r}
# clean data

df$ObservationDate <- as.character(df$ObservationDate)
df <- df[str_detect(df$ObservationDate, "-"),]
df$ObservationDate=format(as.Date(df$ObservationDate),"%Y-%m")
df <- df[complete.cases(df[,7]),]
```




EDA
=======================================================================



Column {data-width=400}
-------------------------------------
- To find the woking range of each equipments, a plot of five main equipments respect to the depth of water of coral found is made. 

- From the plot, it is clearly that most corals are found between 0m - 1000m. 

- Among all devices, the ROV has the widest range of finding corals. 

- Besides, most corals found in deep sea are detected by the ROV.

```{r}
# clean data
violin<-df[,c(10,18)]
violin <- na.omit(violin)
#summary(violin)

violin = filter(violin, SamplingEquipment == c('ROV','submersible','trawl','towed camera', 'longline'))
```


### Sea coral found by 5 main equipments with depth
```{r fig.height=5, fig.width=8}
ggplot(violin, aes(x = SamplingEquipment, y = DepthInMeters)) +
     geom_boxplot((aes(fill = factor(SamplingEquipment))),
                outlier.size = 2, 
                colour = "black",
                na.rm = TRUE,
                alpha = 0.2,
                width = 0.55,
                outlier.shape = NA,
                show.legend = FALSE) + 
    geom_violin((aes(fill = factor(SamplingEquipment))),
                color = NA,
                alpha = 0.5,
                na.rm = TRUE,
                width = 1.3,
                show.legend = FALSE)+
  coord_cartesian(ylim = c(-500, 3000))+
  ggtitle("Sea coral found by 5 main equipments with depth")
```




```{r}
# clean data
violin_coral<-df[,c(4,10)]
violin_coral <- na.omit(violin_coral)
#summary(violin_coral)

violin_coral = filter(violin_coral, VernacularNameCategory == c('gorgonian coral','sponge (unspecified)','sea pen','glass sponge', 'soft coral','demosponge'))
```








Column {data-width=400}
-------------------------------------
- The plot below shows the depth of main categories of corals founded. 

- From the plot, we can find that eccept demosponge which are found near sea level, all other are mainly found between 400m- 1500m. 

- The gorgonian coral has the widedst range of distribution while the demosponge has the shortest.


### Sea coral found by main types with depth

```{r fig.height=5, fig.width=8 }
ggplot(violin_coral, aes(x = VernacularNameCategory, y = DepthInMeters)) +
     geom_boxplot((aes(fill = factor(VernacularNameCategory))),
                outlier.size = 2, 
                colour = "black",
                na.rm = TRUE,
                alpha = 0.2,
                width = 0.5,
                outlier.shape = NA,
                show.legend = FALSE) + 
    geom_violin((aes(fill = factor(VernacularNameCategory))),
                color = NA,
                alpha = 0.55,
                na.rm = TRUE,
                width = 1.3,
                show.legend = FALSE)+
  coord_cartesian(ylim = c(-500, 3000))+
  ggtitle("Sea coral found by main types with depth")
```


```{r}
temp <- data.frame(df$VernacularNameCategory,0)
temp <- temp %>%
  group_by(df.VernacularNameCategory) %>%
  count(df.VernacularNameCategory)
temp <- temp[order(-temp$n),]
names(temp)<- c('Category','Counts')
temp<-na.omit(temp)
```


Column {data-width=400}
-------------------------------------
- Plot below shows the founded number of each corals. 

- The gorgonian coral has the largest number among all corals.

- Most corals are gorgonian coral,sponge (unspecified),sea pen,glass sponge, soft coral and demosponge, and some of them are showed in previous image.

### World Sea Coral number
```{r}

temp %>%
  ggplot( aes(x=reorder(Category,Counts), y=Counts)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
     xlab("Coral Category") +
    theme_bw()+
    ggtitle("World Sea Coral number")

```






Sea Coral Map
=======================================================================
- The plot is a world map of position of main coral categories.

- Most corals are found near the coast espacially the West and East coast of North Amarican.

- Few corals are found Indian Ocean.

- The coast of Alaska contains huge amount of the different corals espacially the soft corals.

- Gorgonian coral which is the greatest in number has the most diverse distribution arong all the world.

## Geo plot of different sea coral

```{r}
#load map data
world <- ne_countries(scale = "medium", returnclass = "sf")

```

```{r}
#change data type to numeric
df$longitude <- as.character(df$longitude)
df$latitude <- as.character(df$latitude)

df$longitude <- as.numeric(df$longitude)
df$latitude <- as.numeric(df$latitude)
```

```{r}
#keep commen category and change others to 'Others'
df_map <- df[,c(4,8,9)]

df_map = filter(df_map, VernacularNameCategory%in%c('gorgonian coral','sponge (unspecified)','sea pen','glass sponge', 'soft coral','demosponge'))

```


```{r fig.height=10, fig.width=20}
#plot geo map
ggplot(data = world) +
    geom_sf(color = "black", fill = "lightblue") +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("World Sea Coral map")+
    geom_point(data = df_map, aes(x = longitude, y = latitude, fill = factor(VernacularNameCategory) ),shape = 23)+
  scale_fill_discrete(name="Coral Categories")
```



```{r}
#set bins
net_df <- df[,c(4,10)]
#summary(net_df$DepthInMeters)
net_df$Depth <- NA
net_df$Depth[net_df$DepthInMeters>=0]<-'0-400'
net_df$Depth[net_df$DepthInMeters>400]<-'401-800'
net_df$Depth[net_df$DepthInMeters>800]<-'801-1200'
net_df$Depth[net_df$DepthInMeters>1200]<-'1201-1600'
net_df$Depth[net_df$DepthInMeters>1600]<-'1601-2000'
net_df$Depth[net_df$DepthInMeters>2000]<-'>2000'
net_df<-net_df[,-2]
names(net_df)=c('Category','Depth')
```



```{r}
# groupby depth and count
net_df$Category<-as.character(net_df$Category)
net_df<-net_df %>% group_by(Category) %>% count(Depth)

net_df = filter(net_df, Category%in%c('gorgonian coral','sponge (unspecified)','sea pen','glass sponge', 'soft coral','demosponge'))
net_df <- na.omit(net_df)

# create index
nodes <- data.frame(name=c(as.character(net_df$Category), as.character(net_df$Depth)) %>%
                      unique())
net_df$IDCategory=match(net_df$Category, nodes$name)-1 
net_df$IDDepth=match(net_df$Depth, nodes$name)-1

```

Sankey plot (network)
=======================================================================
- The plot is an sankey network bwteen categoral of corals and the depth of where they were found.
- The plot can be draged or clicked by mouse.
- Most corals are detected between 0m - 400m below the sea level.
- Only few corals are found below 2,000m

## Main depth of five main sea corals found

```{r fig.height=15, fig.width=10}
sankey<-sankeyNetwork(Links = net_df, Nodes = nodes,
              Source = "IDCategory", Target = "IDDepth",
              Value = "n", NodeID = "name", 
              sinksRight=FALSE, nodeWidth=40, fontSize=10, nodePadding=5, height = 400, width = 600
              )

sankey <- htmlwidgets::prependContent(sankey, htmltools::tags$h1("Main depth of five main sea corals found"))

sankey
```

Time Series
=======================================================================

- The plot shows the time series of corals found between 1995 and 2016, where red line is the average and organge tendency.

- We can see that there is an increase of amount around 2016.

- This time series is kind of flat which means the speed of finding new corals is about constant among those years.

## Time Series of Sea Coral from 1990 - 2016
```{r fig.height=5, fig.width=8}

ts <- data.frame(Time=as.character(df$ObservationDate),counts=0)
#format it
ts$Time <- interaction( ts$Time, '-01', sep = "")


# find the number of each month
ts <- ts %>%
  group_by(Time) %>%
  count(Time)

ts$Time=as.Date(ts$Time)

```



```{r fig.height=5, fig.width=8}
ggplot(ts, aes(x=Time, y=n)) +
  geom_line(color="#69b3a2") + 
  xlab("Year")+
  ylab('Number of Coral found')+
  theme_ipsum()+
  ggtitle('Time Series of Sea Coral from 1990')+
  scale_x_date(limits = c(as.Date(2000-03-01,origin = "1990-01-01"),NA))+
   stat_smooth(
  color = "#FC4E07", fill = "#FC4E07",
  method = "loess"
  )

  


```


Word Cloud
=======================================================================

- The plot is a word cloud shows the repository of corals.(Refresh may be needed)
  
- Suprisingly, the most common word is "research" which is an academic word.

- From the word cloud, we can see that most corals are found by those research team, institude and universities.

- Besides, finding a new sea corals depends more on science rather than economy or business.

```{r}
##Create a vector containing only the text
text <- df$Repository
# Create a corpus  
docs <- Corpus(VectorSource(text))
```

```{r}
#remove stop words and others
docs <- docs %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs <- tm_map(docs, content_transformer(tolower))
docs <- tm_map(docs, removeWords, stopwords("english"))
```


## Word Cloud
```{r fig.height=5, fig.width=8}
dtm <- TermDocumentMatrix(docs) 
matrix <- as.matrix(dtm) 
words <- sort(rowSums(matrix),decreasing=TRUE) 
df_wc <- data.frame(word = names(words),freq=words)

# 
# wordcloud(words = df_wc$word, freq = df_wc$freq, min.freq = 1,
#            max.words=150, random.order=FALSE, rot.per=0.35,
#            colors=brewer.pal(8, "Dark2"))
# 

wordcloud2(df_wc,size=0.8)

```


